---
title: "Reptile Research output"
author: ""
date: "2024-01-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

loading all the packages we will need

```{r}
library(dismo)
library(rworldmap)
library(sf)
library(geodata) 
library(data.table)
library(dplyr)
library(ggplot2)
library(glmmTMB)
library(DHARMa)
library(bbmle)
library(janitor)
library(phylobase) 
library(maps)
library(lme4)
library(arm)
library(ggfortify)
library(pscl)
library(gridExtra)
library(tidyverse)
library(phylosignal)
library(Matrix)
library(multcomp)
library(MuMIn)
library(car)
library(usdm)
```

set our working directory and call our data set

```{r}
setwd("")
AllReptiles <- read.csv("All_Reptiles.csv")
unique(AllReptiles$Suborder)
```

now we have the data lets look at which

```{r}
 sum(AllReptiles$Npapers3rdApproach) #total number of papers

OnIsland <-  subset(AllReptiles, Inisland_reptiles == "Yes")
OnIsland
sum(OnIsland$Npapers3rdApproach)

OffIsland <- subset(AllReptiles, Inisland_reptiles == "No")

top10 <- dplyr::top_n(OnIsland, 10, Npapers3rdApproach); sum(top10$Npapers3rdApproach); rm(top10)
top10 <- dplyr::top_n(OffIsland, 10, Npapers3rdApproach); sum(top10$Npapers3rdApproach); rm(top10)

island_papers <- AllReptiles[AllReptiles$Inisland_reptiles == "Yes", "Npapers1stApproach"]
mainland_papers <- AllReptiles[AllReptiles$Inisland_reptiles == "No", "Npapers1stApproach"]

average_papers <- wilcox.test(island_papers, mainland_papers)


```

90617 total papers

2438 island restricted reptile species contain a total of 6586 papers the top 10 most studied species account for 1712 of these papers or 26%. The mean number of papers per species is 2.7

out of the 8094 non island restricted reptiles 84031 papers are published on them, the top 10 most studied species account for 13463 of this or 16% of papers. The mean number of papers per species is 10.4

```{r}
length(which(OnIsland$Npapers1stApproach==0)) 
length(which(OffIsland$Npapers1stApproach==0)) 

length(which(OnIsland$Npapers1stApproach<=5))
```

1161 island restricted species have no publications, this is 47.6% of species.

A vast majority (2209 or 90.6%) of island species have less than 5 studies

```{r}
AllReptiles <- AllReptiles %>%
   mutate(tenth_log = log(Npapers3rdApproach, 10)) #taking the log10 of the papers so they can fit on graph easier

point_colors <- c("Yes" = "darkblue", "No" = "red")
point_alpha <- c("Yes" = 0.5, "No" = 0.4)


  fig1 <- ggplot(AllReptiles, aes(x = Year, y = Npapers1stApproach, label = Species, color = Inisland_reptiles)) +
  geom_jitter(aes(alpha = ifelse(Inisland_reptiles == "Yes", 0.7, 0.2)), size = 2, fill = "white", shape = 19) +  # Set shape to 21 for better visibility
  geom_text(aes(label = ifelse(Npapers1stApproach > 800, as.character(Species), '')), size = 3, hjust = -0.05, vjust = 0.05) +
  labs(title = "Number of Papers Published on Reptile Species Over Time",
       x = "Year of Description", y = "Number of Resaerch Outputs",
       color = "Island-restricted Reptile") +
  theme_minimal(base_size = 10) +
  theme(axis.title = element_text(face = "bold"),
        axis.text = element_text(size = 8),
        axis.line = element_line(colour = "black")) +
  scale_color_manual(values = point_colors) +  # Set custom point colors
  scale_alpha_continuous(range = c(0.2, 0.7), guide = "none")  # Hide alpha legend


print(fig1)



ggplot(AllReptiles, aes(x = Year, y = Npapers1stApproach)) +
  geom_point(alpha = 0.3, color = "blue") +  # Reduce opacity and change color
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add a linear trend line
  labs(title = "Number of Papers Published on Reptile Species Over Time",
       x = "Publication Year",
       y = "Number of Papers",
       color = "Island-restricted reptile") +
  scale_x_continuous(labels = NULL) +  # Format x-axis labels with commas
  scale_y_continuous(labels = NULL) +  # Format y-axis labels with commas
  theme_minimal()  # Apply a minimal theme


##remake the onIslands dataset with the new log10

OnIsland <-  subset(AllReptiles, Inisland_reptiles == "Yes")

#now we make one just for islands

(fig2 <-  ggplot(OnIsland, aes(x = Year, y = Npapers1stApproach, label = Species)) +
   geom_point(alpha = 0.7, size = 1) +
   geom_text(aes(label = ifelse(Npapers1stApproach > 100, as.character(Species), '')),
             size = 3, hjust = -0.05, vjust = 0.05) +
   labs(x = NULL, y = "Number of papers") +
   theme(panel.grid.minor = element_blank(),
         panel.grid.major = element_blank(),
         panel.background = element_blank(),
         axis.title = element_text(size = 7, face = "bold"),
         axis.line = element_line(colour = "black"),
         axis.text = element_text(size = 6, colour = 'black')))


 
```

Is biogeographic realm a factor?

```{r}
#for islands

(mapon <- ggplot(OnIsland[!is.na(OnIsland$MainBiogRealm), ], 
                 aes(x = log2(Npapers3rdApproach+1)))+
    geom_histogram(fill='grey50', color='grey50')+
    scale_fill_brewer(type = 'qual', palette = 2)+
    labs(title = "Distribution of research output across biogeographic realms on islands",
    x=expression(Log[2]~ "number of papers"), y="Count")+
    theme_bw()+
    theme(legend.title = element_blank(),
          legend.position = 'none')+
    facet_wrap(~MainBiogRealm, nrow = 4, ncol = 3, scales = "free_y"))



#for mainland
(mapoff <- ggplot(OffIsland[!is.na(OffIsland$MainBiogRealm), ], 
                 aes(x = log2(Npapers3rdApproach+1)))+
    geom_histogram(fill='grey50', color='grey50')+
    scale_fill_brewer(type = 'qual', palette = 2)+
    labs(title = "Distribution of research output across biogeographic realms on mainlands",
    x=expression(Log[2]~ "number of papers"), y="Count")+
    theme_bw()+
    theme(legend.title = element_blank(),
          legend.position = 'none')+
    facet_wrap(~MainBiogRealm, nrow = 4, ncol = 3, scales = "free_y"))

grid.arrange(mapon, mapoff, ncol = 2)



```

As we would expect more studied species across all realms

Does IUCN redlist category play a role in the number of papers published on a species

```{r}
right_order <- c("DD", "LC", "NT", "VU", "EN", "CR", "EW") #Making sure the categories are in the correct order for the graph, removing the extinct category due to outliers and not being relevent 
AllReptiles$ThreatStatus <- factor(AllReptiles$ThreatStatus, levels = right_order)


OnIsland <-  subset(AllReptiles, Inisland_reptiles == "Yes", ThreatStatus != "EX") #have to redefine on and off island so the right order goes through
OffIsland <- subset(AllReptiles, Inisland_reptiles == "No", ThreatStatus != "EX")
OnIsland

#MAINLAND

IUCN_bar_mainland <- ggplot(OffIsland, aes(x = ThreatStatus, y = AvgStudiesPerSpecies_Mainland, fill = ThreatStatus)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_hline(yintercept = 11.678, linetype = "dashed", color = "red", size = 1) + #means not including NA or extinct species
  labs(title = "Average Research Outputs by IUCN Category the mainland",
       x = " IUCN Categories",
       y = "Average research output per species") +
  theme_minimal()

IUCN_bar_mainland


#ISLAND

IUCN_bar_island <- ggplot(AllReptiles, aes(x = ThreatStatus, y = AvgStudiesPerSpecies_Island, fill = ThreatStatus)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_hline(yintercept = 2.154, linetype = "dashed", color = "red", size = 1) +
  labs(title = "Average Research Outputs by IUCN Category on islands",
       x = " IUCN Categories",
       y = "Average research output per species") +
  theme_minimal()

IUCN_bar_island
grid.arrange(IUCN_bar_island, IUCN_bar_mainland, ncol = 2) #plotting them together


#########one with them side by side 
#need a new dataframe 

right_order2 <- c("DD_Island", "DD_Mainland", "LC_Island", "LC_Mainland", "NT_Island", "NT_Mainland", "VU_Island", "VU_Mainland", "EN_Island", "EN_Mainland", "CR_Island", "CR_Mainland", "EW_Island", "EW_Mainland") #Making sure the categories are in the correct order for the graph, removing the extinct category due to outliers and not being relevent 

sum(is.na(AllReptiles$ThreatStatus_Plus))

     AllReptiles$ThreatStatus_Plus <- factor(AllReptiles$ThreatStatus_Plus, levels = right_order2)


averages <- aggregate(Npapers1stApproach ~ ThreatStatus + Inisland_reptiles, data = AllReptiles, FUN = mean)
     
     

ggplot(AllReptiles, aes(x = ThreatStatus, y = Npapers1stApproach, fill = Inisland_reptiles)) +
  stat_summary(fun = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = position_dodge(width = 0.9), width = 0.1) +
  labs(title = "Average Research Output Per Species by IUCN Category",
       x = "IUCN Category", y = "Average Research Output",
       fill = "Island Restricted") +
  scale_fill_manual(values = c("Yes" = "gray40", "No" = "lightgrey")) +
  scale_x_discrete(labels = c("NA" = "NC")) + 
  theme_minimal()

#cant do boxplots as the data is just average papers per category and also having some issues with getting them to plot in the right order



IUCN_boxplot <- ggplot(AllReptiles, aes(x = ThreatStatus, y = Npapers1stApproach, fill = Inisland_reptiles)) +
  geom_boxplot() +
  labs(title = "Average Research",
       x = "IUCN Categories",
       y = "Average research output per threat category") +
  theme_minimal()

IUCN_boxplot #looks pretty bad
```

We can see from this graph that the grand mean for papers published is far higher for mainland species which is what we established previously but here we can see that within IUCN categories some are far more studied than others. Species that are extinct in the wild are not studied on islands at all and below th mean for the mainland populations and vulnerable species on the mainland are far more likely to be studied compared to their island counterparts as they are far above the mainland mean where as island species are only sightly above

Running an Mixed effect models

```{r}

OnIsland_NoBlanks <- OnIsland[complete.cases(OnIsland$ThreatStatus), ] #removing the blank NA species from threat status so we can run the LM
OffIsland_NoBlanks <- OffIsland[complete.cases(OffIsland$ThreatStatus), ]

AllReptiles_NoBlank <- AllReptiles[complete.cases(AllReptiles$ThreatStatus), ]

sum(!complete.cases(AllReptiles_NoBlank$ThreatStatus))


lm1 <- lm(Npapers3rdApproach ~ Inisland_reptiles * ThreatStatus, data = AllReptiles_NoBlank)
summary(lm1)
lm1

# above we ran a ancova to study the differences in average number of studies while controlling for the affect of IUCN categories. Average_studies_per_IUCN is the average number of studies for species in each IUCN category and is our response variable. InIsland_reptiles is telling us if  the species is found on islands or not and is our predictor variable. ThreatStatus tells us the species IUCN status and is our categorical variable. This shows that after accounting for covariables there is a statistacally significant difference between the number of studies on the island and mainlan for each IUCN catagory

t_test <- t.test(Npapers3rdApproach ~ Inisland_reptiles, data = AllReptiles_NoBlank, var.equal = FALSE)

t_test

IUCN_glm <- glm(Npapers3rdApproach ~ ThreatStatus * Inisland_reptiles, data = AllReptiles_NoBlank, family = poisson(link = log))
summary(IUCN_glm)
IUCN_glm

#another test using Welch's t-test which is robust to unequal varaince between groups also results in a significant result accepting my hypothesis 

Kruskal <- kruskal.test(Avg_studies_per_IUCN ~ Inisland_reptiles, data = AllReptiles_NoBlank)
Kruskal

IUCN_Kruskal <- kruskal.test(Avg_studies_per_IUCN ~ ThreatStatus, data = AllReptiles)
IUCN_Kruskal

confint(lm1)

lm_intercept <- coef(lm1)[1]
lm_slope <- coef(lm1)[2]


```

GLMM and posthoc on IUCN status data

```{r}
#adding new collum which combines island mainland data and iucn status
AllReptiles <- AllReptiles %>%
 mutate(ThreatStatus_Plus = ifelse(ThreatStatus == "", "NA", paste(ThreatStatus, ifelse(Inisland_reptiles == "Yes", "_Island", "_Mainland"), sep = "_")))
AllReptiles

#now run the GLMM

GLMMIUCN_NB <- glmer(Npapers1stApproach ~ ThreatStatus_Plus + (1|Family), data = AllReptiles, family  = negative.binomial(theta = 1)) 


simulateResiduals(fittedModel = GLMMIUCN_NB, plot = T, n = 1000) 
testDispersion(GLMMIUCN_NB, plot = F, type = 'PearsonChisq', alternative = 'greater') #did not fit well with the negative binomial but we dont have many other options

IUCN_output<-summary(GLMMIUCN_NB)



GLMMrealm_coef <- fixef(GLMMIUCN_NB)

GLMMIUCN_output <- capture.output(summary(GLMMIUCN_NB))
GLMMIUCN_file <- "glmm_iucn.txt"
writeLines(GLMMIUCN_output, GLMMIUCN_file)

### POST HOC P VALUE COMPARISON



IUCN_MCP <- glht(GLMMIUCN_NB, linfct = mcp(ThreatStatus_Plus = "Tukey"))

summary(IUCN_MCP)
IUCN_summary <- capture.output(summary(IUCN_MCP))

# Define the file path for saving the text file
IUCNMC_file <- "multiple_comparison_results.csv"

# Write the information to a text file
write.csv(IUCN_summary, IUCNMC_file)
```

all tests even when data is not normally distributed and this is taken account for in the Kruskal Wallace test we can see that there is a significantly different number of studies between island and mainland species of the same IUCN categories.

Graphing data to see if the bio geographic realm is a factor in how reptile species are researched

```{r}
realm_study_spread <-
  ggplot(data = AllReptiles, aes(x = MainBiogRealm, y = tenth_log)) +
  facet_wrap(~Inisland_reptiles) +
  geom_boxplot() +
  geom_smooth(formula = y ~ x, method = lm) +
  labs(x = "Realm", y = "Studies (log2 scale)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
realm_study_spread

filtered_data<- AllReptiles %>% filter(RealmPlus != "NA")

filtered_data <- filtered_data %>%
  group_by(RealmPlus) %>%
  mutate(AvgPapersPerSpeciesRealm = sum(Npapers1stApproach) / n())

Realm_Bar <- ggplot(filtered_data, aes(x = RealmPlus, y = AvgPapersPerSpeciesRealm, fill = RealmPlus)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Research",
       x = "Biogeogrpahic Realm ",
       y = "Average Research Output") +
  theme_minimal()


Realm_Bar



realmplus_study_spread <-
  ggplot(data = filtered_data, aes(x = RealmPlus, y = tenth_log)) +
  geom_boxplot() +
  geom_smooth(formula = y ~ x, method = lm) +
  labs(x = "Biogeographic Realm", y = "Studies (log10)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

realmplus_study_spread


ggplot(filtered_data, aes(x = MainBiogRealm, y = Npapers1stApproach, fill = Inisland_reptiles)) +
  stat_summary(fun = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = position_dodge(width = 0.9), width = 0.1) +
  labs(title = "Average Research Output Per Species by Biogeographic Realm",
       x = "Biogeographic Realm", y = "Average Research Output",
       fill = "Island Restricted") +
  scale_fill_manual(values = c("Yes" = "gray40", "No" = "lightgrey")) +
  scale_x_discrete(labels = c("Australasi" = "Australasia")) +
  theme_minimal()




Realm_Kruskal <- kruskal.test(Npapers1stApproach ~ MainBiogRealm, data = AllReptiles)
Realm_Kruskal

Realm_lm <- lm(Npapers3rdApproach ~ Inisland_reptiles * MainBiogRealm, data = AllReptiles_NoBlank)
summary(Realm_lm)


glm_Realm <- glm(Npapers3rdApproach ~ MainBiogRealm * Inisland_reptiles, data = AllReptiles_NoBlank, family = poisson(link = log))
summary(glm_Realm)
glm_Realm


```

here we run various models to accurately model the relationship between the number of papers published on a species and the bio geographic realm they are found in for islands and the mainland. We are trying to predict how the number of papers changes based on the realm. The categorical predictor is the MainBiogRealm categorical variable. Another predictor is Inisland_reptiles which tells us if the species is found on islands or mainland as we expect this to influence the number of papers published on a species. Its inclusion in the model allows us to assess how the average count of papers differs between island and mainland species. As the interaction term of Inisland_reptiles is significant at across most of the regions it suggests the effect of the bio geographic realm on the number of papers published depends on whether the species is found on the island or the mainland. To conclude this model suggests that the average number of studies is significantly influenced by the bio geographical realm, whether the reptile is on an island, and the interaction between these factors.

we can also interpret the coefficients which tell us compared to the reference intercept an island reptile has an overall negative impact on the log odds meaning being on an island has negative impact on the number of papers that would be published on the species. The largest change in log odds between island and mainland is in the Neartic realm. This large negative value of -2.51341 indicates that being on an island in the Neartic realm is associated with a decrease in the number of studies.

We now want to run a GLMM for the same premise of biogeographic realms being different between islands and mainlands

```{r}
#we need to combine the data on realm and if they are found on islands or mainlands for the GLMM

AllReptiles <- AllReptiles %>%
 mutate(RealmPlus = ifelse(MainBiogRealm == "", "NA", paste(MainBiogRealm, ifelse(Inisland_reptiles == "Yes", "_Island", "_Mainland"), sep = "_")))
AllReptiles

#then turn them into intgers



# AllReptiles$RealmPlus <- as.integer(as.character(factor(AllReptiles$RealmPlus,labels = c(NA,14,13,12,11,10,9,8,7,6,5,4,1,2,3))))
#    AllReptiles$IUCN <- scale(AllReptiles$RealmPlus, center = T, scale = T) # z-transform variable to make it comparable

#now run the GLMM

GLMMrealm_NB <- glmer(Npapers1stApproach ~ RealmPlus + (1|Family), data = AllReptiles, family  = negative.binomial(theta = 1)) 


simulateResiduals(fittedModel = GLMMrealm_NB, plot = T, n = 1000) 
testDispersion(GLMMrealm_NB, plot = F, type = 'PearsonChisq', alternative = 'greater') #did not fit well with the negative binomial but we dont have many other options

summary(GLMMrealm_NB)

GLMMrealm_coef <- fixef(GLMMrealm_NB)

### POST HOC P VALUE COMPARISON

Realm_MCP <- glht(GLMMrealm_NB, linfct = mcp(RealmPlus = "Tukey"))

summary(Realm_MCP)

#Do we need to correct this for size of the realm? like dividing the number of studies by the size of the realm


```

```{r}

# Calculate average papers per family
average_papers_per_family <- AllReptiles %>%
  group_by(Suborder, Family) %>%
  summarize(AveragePapers = mean(Npapers1stApproach, na.rm = TRUE)) %>%
  mutate(Family = reorder(Family, -AveragePapers)) %>%
  arrange(Suborder, -AveragePapers)

#can also arrange in random order but still in grouped into orders 
#################################################################################
#average_papers_per_family <- AllReptiles %>%
  #group_by(Order, Family) %>%
  #summarize(AveragePapers = mean(Npapers3rdApproach, na.rm = TRUE)) %>%
  #mutate(Family = factor(Family, levels = unique(Family)))
#################################################################################


# Create empty bars for spacing
empty_bar <- 20
to_add <- data.frame(matrix(NA, empty_bar * nlevels(average_papers_per_family$Suborder), ncol(average_papers_per_family)))
colnames(to_add) <- colnames(average_papers_per_family)
to_add$Suborder <- as.character(to_add$Family)
to_add$Family <- rep(levels(average_papers_per_family$Suborder), each = empty_bar)
average_papers_per_family <- rbind(average_papers_per_family, to_add)
average_papers_per_family <- average_papers_per_family %>% arrange(Family, Suborder)
average_papers_per_family$id <- seq(1, nrow(average_papers_per_family))

# Get the name and the y position of each label
label_data <- average_papers_per_family
number_of_bar <- nrow(label_data)
angle <- 90 - 360 * (label_data$id - 0.5) / number_of_bar
label_data$hjust <- ifelse(angle < -90, 1, 0)
label_data$angle <- ifelse(angle < -90, angle + 180, angle)

# Create a circular bar plot
p <- ggplot(average_papers_per_family, aes(x = as.factor(id), y = AveragePapers, fill = Suborder)) +
  geom_bar(stat = "identity", alpha = 0.5) +
  ylim(-100, 120) +
  theme_minimal() +
  theme(
    legend.position = "right",
    axis.text = element_blank(),
    axis.title = element_blank(),
    panel.grid = element_blank(),
    plot.margin = unit(rep(-1, 4), "cm")
  ) +
  scale_y_continuous(trans = "log2") +
  coord_polar(start = 0) +
  geom_text(data = label_data, aes(x = id, y = AveragePapers + 10, label = paste(Family, round(AveragePapers, 2)), hjust = hjust),
            color = "black", fontface = "bold", alpha = 0.6, size = 2.5, angle = label_data$angle, inherit.aes = FALSE)+
  labs(fill = "Suborder")

# Display the plot
print(p)

```

new table on what famalies are least studied

```{r}
unique_families <- unique(AllReptiles$Family)

family_names <- character()
zero_studies_proportion <- numeric()
species_count <- numeric()
island_proportion <- numeric()
total_species_count <- numeric()

for (family in unique_families) {
  # Subset the data for the current family
  family_data <- subset(AllReptiles, Family == family)
  zero_studies_count <- sum(family_data$Npapers1stApproach == 0) #find how many have 0 studies
  total_species <- nrow(family_data)
  island_species <- sum(family_data$Inisland_reptiles == "Yes") #find the proportions for these
  #make output into vectors 
  family_names <- c(family_names, family)
  zero_studies_proportion <- c(zero_studies_proportion, zero_studies_count / total_species)
  species_count <- c(species_count, total_species)
  island_proportion <- c(island_proportion, island_species / total_species)
  total_species_count <- c(total_species_count, total_species)
  
}

#new data frame
study_rank <- data.frame(
  Family = family_names,
  ZeroStudiesCount = zero_studies_proportion * species_count,
  ZeroStudiesPercentage = 100 * zero_studies_proportion,
  PercentageIslandSpecies = 100 * island_proportion,
  SpeciesCount = species_count
  
)

least_studied <- study_rank[order(study_rank$ZeroStudiesPercentage, decreasing = TRUE), ]

# Select the top 10 families
ten_least_studied <- head(least_studied, 10)

# Print the result
print(ten_least_studied)
write.csv(ten_least_studied, file = "Least_Studied_Famalies.csv", row.names = FALSE)

####Also for the most studied Famalies

family_summary <- AllReptiles %>%
  group_by(Family) %>%
  summarize(
    MeanStudies = mean(Npapers1stApproach, na.rm = TRUE),
    NumSpecies = n(),
    PercentOnIslands = sum(Inisland_reptiles == "Yes") / n() * 100,
    TotalPapers = sum(Npapers1stApproach, na.rm = TRUE)
  )

ten_most_studied <- family_summary[order(-family_summary$MeanStudies), ] %>% head(10)

write.csv(ten_most_studied, file = "Most_Studied_Famalies.csv", row.names = FALSE)

```

```{r}
kruskal_family <- kruskal.test(AveragePapers ~ Family, data = average_papers_per_family)
kruskal_family
```

standarsing data for a GLMM of all factors

```{r}
e1071::skewness(AllReptiles$BodyMass, na.rm = T); e1071::kurtosis(AllReptiles$BodyMass, na.rm = T) #very high kurtosis means sharp peak and long tail to data and also slight right skew
#conclusion for all data is that it should be log 10 transformed 

for (i in 1) {
  AllReptiles$StdYear <- scale(AllReptiles$Year, center = T, scale = T)
  AllReptiles$LogBodyMass <- scale(log10(AllReptiles$BodyMass), center = T, scale = T)
  AllReptiles$LogRangeSize <- scale(log10(AllReptiles$RangeSize), center = T, scale = T)
  AllReptiles$LogElevation <- scale(log10(AllReptiles$Elevation+1), center = T, scale = T)
  AllReptiles$LogRangeRarity <- scale(log10(AllReptiles$RangeRarity), center = T, scale = T)
  AllReptiles$LogNumBioInst <- scale(log10(AllReptiles$TotBiodInst+1), center = T, scale = T)
  
}


summary(AllReptiles$ThreatStatus)
unique(AllReptiles$ThreatStatus) 

AllReptiles$IUCN <- as.integer(as.character(factor(AllReptiles$ThreatStatus,labels = c(5,NA,4,NA,1,2,3))))
AllReptiles$IUCN <- scale(AllReptiles$IUCN, center = T, scale = T) # z-transform variable to make it comparable

#check for multicollinearity
vars <- AllReptiles[ , 39:38]
usdm::vif(vars); rm(vars)


levels(AllReptiles$MainBiogRealm)

levels(AllReptiles$MainBiogRealm) # Australasia is missing an 'a'; fixing it next 
AllReptiles$MainBiogRealm <- factor(AllReptiles$MainBiogRealm,
                               labels = c(" ", "Afrotropic","Australasia","IndoMalay","Neartic","Neotropic","Oceania","Paleartic"))


colnames(AllReptiles)

dataset <- AllReptiles%>%
  dplyr::select(Species, Npapers1stApproach, Family, MainBiogRealm, StdYear:IUCN)
summary(dataset)

dataset <- droplevels(dataset[complete.cases(dataset) , ])

dataset                      
                      
```

Running different GLMMs

```{r}
GLMM1 <- glmer(Npapers1stApproach ~ LogBodyMass + LogRangeSize + (1|Family), data = dataset, family = poisson(link = log))

simulateResiduals(fittedModel = GLMM1, plot = T, n = 1000) 
testDispersion(GLMM1, plot = F, type = 'PearsonChisq', alternative = 'greater') #plotting resiuduals and we can see that it is not a great fit 


GLMM1_NB <- glmer(Npapers1stApproach ~ LogBodyMass + LogRangeSize + (1|Family), data = dataset, family = negative.binomial(theta = 1))  #theta is the dispersion value
GLMM1_NB # suggests increase in bodymass or range size is associated in an increase in the number of papers publised on that species 

simulateResiduals(fittedModel = GLMM1_NB, plot = T, n = 1000) 
testDispersion(GLMM1_NB, plot = F, type = 'PearsonChisq', alternative = 'greater') #negative binomial residuals plot better 

GLMM2 <- glmer(Npapers1stApproach ~ LogElevation + LogRangeRarity + (1|Family), data = dataset, family =   poisson(link = log))
GLMM2
simulateResiduals(fittedModel = GLMM2, plot = T, n = 1000) 
testDispersion(GLMM2, plot = F, type = 'PearsonChisq', alternative = 'greater')

GLMM2_NB <- glmer(Npapers1stApproach ~ LogElevation + LogRangeRarity + (1|Family), data = dataset, family  = negative.binomial(theta = 1))
GLMM2_NB # suggests elevation and range rarity both have a negative impact on papers published on a species 
simulateResiduals(fittedModel = GLMM2_NB, plot = T, n = 1000) 
testDispersion(GLMM2_NB, plot = F, type = 'PearsonChisq', alternative = 'greater') #only slightly better this time for fitting the residuals

GLMMALL <- glmer(Npapers1stApproach ~ LogElevation + LogRangeRarity + StdYear + LogBodyMass + LogNumBioInst + IUCN + LogRangeSize + (1|Family), data = dataset, family  = negative.binomial(theta = 1))#global GLMM

simulateResiduals(fittedModel = GLMMALL, plot = T, n = 1000) 
testDispersion(GLMMALL, plot = F, type = 'PearsonChisq', alternative = 'greater') 

summary(GLMMALL)

save(GLMMALL, file = 'nbGLMM_global.Rdata')

rm(list=setdiff(ls(), c("mydata", "dataset", "nbmod")))
load("nbGLMM_global.Rdata")
ranef(GLMMALL) #intercepts for familys, showing the intercept which can be plotted to show which families are affected by these factors more than others


fixef(GLMMALL) #what we need for making the dot plot on the direction of the different affects on the number of studies

#lattice package can also be used to plot slopes for the random or fixed affects

#circular bar plot can also plot the intercept for the random effects of the families. same as the one made before but just looking at intercept instead
global_model <- update(GLMMALL, na.action = na.fail)

#######Using model selection to find the most optimal model for the GLMM

dredge(global_model)

```

making a plot for the GLMM

```{r}
#splitting the main dataset into island and mainland for the GLMM

###Island Reptiles 


dataset_ONI <- AllReptiles %>%
  dplyr::filter(Inisland_reptiles == "Yes") %>%
  dplyr::select(Species, Npapers1stApproach, Family, MainBiogRealm, StdYear:IUCN)
summary(dataset_ONI)


dataset_ONI <- droplevels(dataset_ONI[complete.cases(dataset_ONI) , ])

head(dataset_ONI)

#run the GLMM

GLMM_ONI <- glmer(Npapers1stApproach ~ LogElevation + LogRangeRarity + StdYear + LogBodyMass + LogNumBioInst + LogRangeSize + IUCN + (1|Family), data = dataset_ONI, family  = negative.binomial(theta = 1))

simulateResiduals(fittedModel = GLMM_ONI, plot = T, n = 1000) 
testDispersion(GLMM_ONI, plot = F, type = 'PearsonChisq', alternative = 'greater') 

vif(GLMM_ONI) #Checking for multicollinearity in the GLMM values of 5 or more are indicative of that. 

summary(GLMM_ONI, type = "both")
save(GLMM_ONI, file = 'nbGLMM_ON_Islands.Rdata')

performance::r2(GLMM_ONI)

###Finding the most useful model to use
All_ONI<- update(GLMM_ONI, na.action = na.fail)

d1 <- dredge(All_ONI,  m.lim = c(6,7)) #the best model actually didnt include IUCN as a factor but the next best model did with little difference so Will inculude it most likley
#  save(d1, file = 'GLMMONI_limited_dredge.Rdata') # so that we dont have to rerun i 

load("GLMMONI_limited_dredge.Rdata")
d1[1,] #the most usefull model we will use for modle averaging 

load("nbGLMM_ON_Islands.Rdata")

###Mainland Reptiles

dataset_OFFI <- AllReptiles %>%
  dplyr::filter(Inisland_reptiles == "No") %>%
  dplyr::select(Species, Npapers1stApproach, Family, MainBiogRealm, StdYear:IUCN)
summary(dataset_OFFI)


dataset_OFFI <- droplevels(dataset_OFFI[complete.cases(dataset_OFFI) , ])

dataset_OFFI  


GLMM_OFFI <- glmer(Npapers1stApproach ~ LogElevation + LogRangeRarity + StdYear + LogBodyMass + LogNumBioInst + LogRangeSize + IUCN + (1|Family), data = dataset_OFFI, family  = negative.binomial(theta = 1))

simulateResiduals(fittedModel = GLMM_OFFI, plot = T, n = 1000) 
testDispersion(GLMM_OFFI, plot = F, type = 'PearsonChisq', alternative = 'greater') 

summary(GLMM_OFFI)

performance::r2(GLMM_OFFI)


vif(GLMM_OFFI) #Checking for multicollinearity in the GLMM values of 5 or more are indicative of that. 

###Finding the most usefull model based on AIC
All_OFFI<- update(GLMM_OFFI, na.action = na.fail)

d3 <- dredge(All_OFFI, trace = 2,  m.lim = c(6,7)) #the most useful model for non island reptiles does not include elevation
save(d3, file = 'GLMMOFFI_limited_dredge.Rdata')
d2 #the full list of combinations that takes a very long time to run so we limited it to only models that have 6 or 7 facotrs as we knew the best ones had 6 or 7
save(d2, file = 'GLMMONI_full_dredge.Rdata')

load("GLMMONI_full_dredge.Rdata")
d2

load("GLMMOFFI_limited_dredge.Rdata")
d3[1,]

save(GLMM_OFFI, file = 'nbGLMM_OFF_Islands.Rdata')

###Model Averaging, testing multiple hypothesis and finding the best mix of them 


list(GLMM_ONI)
summary(model.avg(d3, subset = delta < 2)) #use <2 for delta as these are the models that are as good as the top model
summary(model.avg(d1, subset = delta < 2))


AIC(GLMM_OFFI)


###plot fixed effects
load("nbGLMM_ON_Islands.Rdata")
load("nbGLMM_OFF_Islands.Rdata")
coefficients_ONI <- fixef(GLMM_ONI)
coefficients_OFFI <- fixef(GLMM_OFFI)


all_coefficients <- c(coefficients_ONI, coefficients_OFFI)


#Create a dataframe
df_coefficients <- data.frame(
  df_coefficients <- data.frame(
  Factor = rep(names(all_coefficients)),  # Each factor repeats twice for ONI and OFFI
  Coefficient = rep(all_coefficients),    # Coefficients repeat twice for ONI and OFFI
  Dataset = rep(c("Island", "Mainland"), each = length(all_coefficients) / 2),  # Specify the dataset
  Direction = ifelse(rep(all_coefficients,) > 0, "Positive", "Negative")
))



# Print or further analyze df_coefficients
print(df_coefficients)

###Graphing


ggplot(df_coefficients, aes(x = Dataset, y = Factor)) +
  geom_point(aes(size = abs(Coefficient), colour = Direction)) +
  theme_bw() +
  xlab("Island") +
  ylab("Covariate") +
  scale_colour_discrete(name = "Effect direction") +
  scale_size_continuous(name = expression(beta), range = c(1, 10), breaks = seq(-0.05, 0.65, by = 0.1)) +
  theme(
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13),
    axis.text.x = element_text(angle = 90)
  )

```

Now we have a GLMM for all factors we can plot across the different realms

```{r}
formula <- Npapers1stApproach ~ LogElevation + LogRangeRarity + StdYear + LogBodyMass + LogNumBioInst + (1|Family)

levels(dataset$MainBiogRealm)
biorealm <- levels(dataset$MainBiogRealm)
biorealm <- biorealm[-6] #ocean ia removed due to small sample size

outs <- list()

for (i in seq_along(biorealm)) {
  # create subdataset containing only observations from a single biog. realm
  subdataset = droplevels(dataset[dataset$MainBiogRealm==biorealm[i] , ])
  
  # fit the NB mixed model
  nbglmm = glmmTMB(formula = formula, data = subdataset, family = nbinom2)
  
  print(paste0(biorealm[j], " ", performance::r2(nbglmm)))
  
  # residual diagnostics
  pdf(paste("ResidPlot", biorealm[i], '.pdf', sep = '_'))
  simulationOutput = simulateResiduals(fittedModel = nbglmm, plot = T)
  dev.off()
  
  # save model 
  save(nbglmm, file = paste("nbGLMM", biorealm[i], '.Rdata', sep = '_'))
  
  # get result
  res = summary(nbglmm)
  
  # store coefficients
  coefs = as.data.frame(res$coefficients$cond)
  coefs$Predictors <- rownames(coefs) # add predictor names into a new column
  coefs = mutate(coefs, 
                 Realm = paste(biorealm[i]), # add realm names into a new column
                 lower95 = Estimate - 1.96 * `Std. Error`, # get lower CI95%
                 upper95 = Estimate + 1.96 * `Std. Error`) # get upper CI95%
  coefs = coefs[ , c("Realm","Predictors","Estimate","Std. Error","z value","Pr(>|z|)","lower95","upper95")]
  
  outs[[i]] <- coefs
  
  rm(subdataset, nbglmm, simulationOutput, coefs)
  
}
```

research priority index. We hypothesise that Island species would on average rank higher than mainland ones partially due to the lack of studies overall on them

```{r}
Islandspecies <- read.csv("data/Islandspecies_CSV8.csv")


# Assuming "Binomial" and "Species" are the respective common columns
merged_data <- merge(AllReptiles, Islandspecies[, c("Binomial", "ERi")], by.x = "Species", by.y = "Binomial", all.x = TRUE)

# View the merged data
print(merged_data)

### ROi

max_papers <- max(AllReptiles$Npapers1stApproach, na.rm = TRUE)
columns <- c("Species", "Genus", "RangeSize", "ThreatStatus", "Npapers1stApproach", "Inisland_reptiles", "ERi")

research_priority <- merged_data[columns, drop = FALSE]

# Calculate ROi for each species
research_priority <- research_priority %>%
  mutate(
    ROi = 1 - (Npapers1stApproach / max_papers)
  )
  


# Display the resulting data frame
print(research_priority)

### ERi

#need to first give a value to species that are data deficient or not categorised 
#range less than 400000km2 or 33 in the All reptiles database get 0.5. this is for non island restricted reptiles only as we already have the data for IRR and how many islands they reside on is not present in either dataset . Other than this all ERi is based on IUCN status



research_priority <- research_priority %>%
  mutate(
    ERi = ifelse(
      Inisland_reptiles == "No" & ThreatStatus %in% c("VU", "EN", "CR"), 1,
      ifelse(
        Inisland_reptiles == "No" & ThreatStatus %in% c("LC", "NT"), 0,
        ifelse(
          Inisland_reptiles == "No" & ThreatStatus %in% c("DD", "NC") & RangeSize > 33, 0,
          ifelse(
            Inisland_reptiles == "No" & ThreatStatus %in% c("DD", "NC") & RangeSize <= 33, 0.5,
            ERi
          )
        )
      )
    )
  )





print(research_priority)

### CSi


# Calculate the total number of species in each genus
genus_counts <- research_priority %>%
  group_by(Genus) %>%
  summarise(total_species = n())

genus_counts

# Find the most populated genus
most_populated_genus <- genus_counts %>%
  filter(total_species == max(total_species))

most_populated <- sum(most_populated_genus$total_species)

research_priority <- research_priority %>%
  left_join(genus_counts, by = "Genus")
  
research_priority <- research_priority %>%
  mutate(
    CSi = 1 - ((total_species - 1) / (most_populated - 1)))

research_priority = subset(research_priority, select = -c(total_species)) #removing the total species collumn
print(research_priority)

### ROCSi


#find the total number of papers published in reptiles genus not including its own publications

research_priority <- research_priority %>%
  group_by(Genus) %>%
  mutate(total_research_output_within_genus = sum(Npapers1stApproach) - Npapers1stApproach) %>%
  ungroup()


#finding the highest ammount of papers published in any genus
genus_max_research <- research_priority %>%
  group_by(Genus) %>%
  summarize(total_max_research = sum(Npapers1stApproach)) %>%
  ungroup()
  
  
genus_with_highest_total_research <- genus_max_research %>%
  filter(total_max_research == max(total_max_research))
 
most_studied_genus <- sum(genus_with_highest_total_research$total_max_research)


#calculating ROCSi 
research_priority <- research_priority %>%
  mutate(ROCSi = 1 - (total_research_output_within_genus / most_studied_genus))

research_priority = subset(research_priority, select = -c(total_research_output_within_genus)) #removing the total_research_output_within_genus column

print(research_priority)

###RPM

research_priority <- research_priority %>%
  mutate(RPM = ROi + ERi + 0.5  * CSi + 0.5 * ROCSi)

print (research_priority)
write.csv(research_priority, file = "Complete_Reptile_Research_Prioritisation_Metric.csv", row.names = FALSE)

top_15_rpm <- research_priority %>%
  arrange(desc(RPM)) %>%
  head(15)

# Print the top 10 RPM values
print(top_15_rpm)

write.csv(top_15_rpm, file = "Reptile_Research_Prioritisation_Metric.csv", row.names = FALSE)

```

phylogenetic tree code

```{r}
library(ape)
library(cluster) 
library(RColorBrewer)
library(ggplot2)
library(dplyr)
library(tidytree)
library(treeio)
library(data.table)
library(ggtree)
library(plyr)

rt <- "( 
  (Scincoidea, Anguimorpha, Iguania),
  (Lacertiformes, Serpentes, Gekkota),
  (Pleurodira, Cryptodira, Dibamidae),
  (Crocodylia, Rhynchocephalia)
);" #place holder structure of the tree

rttree <- read.tree(text = rt)

# Check the structure of the phylogenetic tree
plot(rttree)

suborder_labs <- rttree$tip.label

x = as_tibble(rttree)


d <- read.csv(choose.files())



# Select relevant columns
columns <- c("Species", "Suborder", "ThreatStatus", "Npapers1stApproach")
d <- d[columns]

# Calculate number of threatened species per suborder
d$threatened <- ifelse(d$ThreatStatus %in% c("VU", "EN", "CR"), 1, 0)
threatened_per_Suborder <- aggregate(threatened ~ Suborder, data = d, sum)

# Merge threatened species counts with the main dataframe
d <- merge(d, threatened_per_Suborder, by = "Suborder", all.x = TRUE)


colnames(d)[colnames(d) == "threatened.x"] <- "IsThreatened"
colnames(d)[colnames(d) == "threatened.y"] <- "threatened"


# Calculate studies per threatened species
d$studspersp <- d$Npapers1stApproach / d$threatened
d$label <- d$Suborder 

d$studspersp[d$studspersp == 0] <- NA


# Join tree data with species data may have to remove suborders with no threatned species
treecon <- full_join(x, d, by = "label")
treecon1 <-  as.treedata(treecon) 

str(treecon1)

breaks <- seq(0,round_any(max(treecon$studspersp,na.rm=TRUE), 10, f = ceiling)  ,length.out=5)
data.limits <- c(0,round_any(max(treecon$studspersp,na.rm=TRUE),10, f = ceiling))
  
ggtree(treecon1, layout="fan", open.angle=45) + 
  geom_tiplab2(hjust=-0.25, size=5.95) + 
  geom_tippoint(aes(fill=studspersp),shape=21,size=15)+
  scale_fill_gradientn(name="Studies per species",breaks=breaks,limits=data.limits,
                       colours=c("white",(brewer.pal(n=9, name="Blues"))[1:9]),
                       na.value="grey50")+
  theme(legend.position=c(0.95,0.45),axis.text = element_text(size=15),
        legend.text = element_text(size=17),
        legend.title = element_text(size=20))+xlim(0,3)



```

